---
title: "Join Lab"
author: "Dr. Shoemaker"
date: "DATA 2401"
output: html_document
---

# Joins and `dplyr` practice

Load the `"nycflights13"` package. You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(tidyverse)
#install.packages("nycflights13")
library(nycflights13)
my_flights <- flights # store my own version
my_flights <- na.omit(my_flights) ## this dataset has data on cancelled flights too, 
                                  ## so we 'll remove those for now

```

Create a dataframe of the average arrival delays for each *destination*

```{r}
avg_arr_delays <- my_flights %>% group_by(dest) %>% summarize(avgArrDelay = mean(arr_delay))
```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}
left_join(avg_arr_delays, airports, join_by(dest == faa))
```

Note here: the column names don't match anymore!

Which airport had the largest average arrival delay?

```{r}
left_join(avg_arr_delays, airports, join_by(dest == faa)) %>% filter(avgArrDelay == max(avgArrDelay))
```

What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each *airline*, then use `left_join()` to join on the "airlines" dataframe

```{r}
avg_dep_delays <- my_flights %>% group_by(carrier) %>% summarize(avgDepDelay = mean(dep_delay))

avg_dep_delays <- left_join(avg_dep_delays, airlines)
```

Which airline had the smallest average departure delay?

```{r}
avg_dep_delay %>% filter(avgDepDelay == min(avgDepDelay))
```

Examples from R4DS:

```{r}

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join:

```{r}
flights %>% 
  semi_join(top_dest)
```

a) What does anti_join(flights, airports, by = c("dest" = "faa")) tell you?

b) What does anti_join(airports, flights, by = c("faa" = "dest")) tell you? Perform the two joins below and consider what the results are.

```{r}
#a) Lists flights to airports not in the faa's airports table, aka international dests.
anti_join(flights, airports, by = c("dest" = "faa"))

#b) Lists airports you cannot flight to directly from NY airports.
anti_join(airports, flights, by = c("faa" = "dest"))
```
